Gender and Mental Depression: Analyzing Census Data Through Statistical Methods

### 2024-10-21 paper/paper.qmd
added marcus's writeup

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(boot)
library(broom.mixed)
library(collapse)
library(dataverse)
library(gutenbergr)
library(janitor)
library(knitr)
library(marginaleffects)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)
# df <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
labeled_df <- read.csv(here::here("data/02-analysis_data/model_data_sample_2k.csv"))
```
# Introduction

@@

```{r}
#| label: fig-education
#| fig-cap: Education
#| warning: false
#| message: false
#| echo: false
labeled_df |>
  ggplot(aes(x = label_education, fill = label_outcome)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(label_sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Mental Depression"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "bottom")
```
## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-lrmodelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))

modelsummary(
  list(
    "Mental Illness Exist" = mental_illness
  ),
  statistic = "mad"
  )
```

@@

# Additional data details

>>>>

---
title: "Gender and Mental Depression: Analyzing Census Data Through Statistical Methods"
author: 
  - Christina Nguyen
  - Marcus Barnes
  - Nibras Ar Rakib
thanks: "Code and data are available at: [https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty](https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(boot)
library(broom.mixed)
library(collapse)
library(dataverse)
library(gutenbergr)
library(janitor)
library(knitr)
library(marginaleffects)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)
library(here)
filtered_df_2018 <- read.csv(here::here("data/02-analysis_data/model_data_2018.csv"))
filtered_df_2021 <- read.csv(here::here("data/02-analysis_data/model_data_2021.csv"))
```

# Introduction

@@

```{r}
#| label: fig-education
#| fig-cap: Education
#| warning: false
#| message: false
#| echo: false
# labeled_df |>
#   ggplot(aes(x = label_education, fill = label_outcome)) +
#   stat_count(position = "dodge") +
#   facet_wrap(facets = vars(label_sex)) +
#   theme_minimal() +
#   labs(
#     x = "Highest education",
#     y = "Number of respondents",
#     fill = "Mental Depression"
#   ) +
#   coord_flip() +
#   scale_fill_brewer(palette = "Set2") +
#   theme(legend.position = "bottom")
```

## Overview
We used the Medical Expenditure Panel Survey dataset from the Integrated Public Use Microdata Series, which contains microdata from different time points related to U.S. healthcare expenditures [@blewett_ipums_2023]. We performed all analysis using a popular statistical language called R [@citeR]. Our dataset contains age, sex, marital status, race, highest level of education attainment, total personal income, and K6 score [@kessler_kessler_2012].All the variable definitions can be found in the Appendix [-@sec-appendix-data-details].

Our goal was to identify people, based on gender and education, who are suffering from non-specific psychological distress based on the Kessler scale [@kessler_kessler_2012] before and after the COVID-19 pandemic.

Our goal was to identify respondents, based on gender and education, who were suffering from non-specific psychological distress based on the Kessler scale before and after the COVID-19 pandemic. To support our analyses, we split into two datasets for 2018 and 2021. We also excluded the respondents whose ages are below 19 and over 80. After enforcing our inclusion criteria, we had 15716 respondents for 2018 and 12685 for 2021. To give a preview of our data, @tbl-2018 and @tbl-2021 show the first five rows for each row for 2018 and 2021, respectively.
```{r}
#| label: fig-education-number-rows
#| fig-cap: Education
#| warning: false
#| message: false
#| echo: false
#| results: hide
print(nrow(filtered_df_2018))
print(nrow(filtered_df_2021))
```
```{r}
#| label: tbl-2018
#| tbl-cap: "Year 2018"
#| warning: false
#| message: false
#| echo: false
kable(filtered_df_2018[1:5,c("year","age","sex","marstat","racea","educ","inctot","k6sum")])
```
```{r}
#| label: tbl-2021
#| tbl-cap: "Year 2021"
#| warning: false
#| message: false
#| echo: false
kable(filtered_df_2021[1:5,c("year","age","sex","marstat","racea","educ","inctot","k6sum")])
```
## Measurement

### Gender and Mental Health

Research consistently shows that women report higher rates of depression than men (CITE). Biological, social, and psychological factors contribute to this disparity. Women may experience unique stressors, such as role strain from balancing work and family obligations, as well as being subject to gender-based discrimination (CITE). These pressures can exacerbate mental health issues, leading to higher rates of depression among women, particularly those in lower socioeconomic or educational strata.

### Educational Attainment and Mental Health

Education is often seen as a protective factor against mental health issues. Studies suggest that higher levels of education correlate with improved mental health outcomes (CITE). Educated individuals tend to have better access to mental health resources, higher socioeconomic status, and greater job security, all of which contribute to lower rates of depression (CITE). Conversely, those with lower education levels may face more financial strain, reduced employment opportunities, and limited access to healthcare, increasing their vulnerability to depression.

### Intersection of Gender, Education, and Depression

Despite the protective nature of education, the gender gap in mental health persists even at higher levels of educational attainment. Some studies suggest that women with high education levels still experience depression due to external stressors, including workplace discrimination and societal expectations (CITE). Additionally, the mental health benefits of education may differ between genders, with men benefiting more from the social status and economic opportunities that education provides, while women may continue to face challenges related to gender roles (cite).

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| label: tbl-lrmodelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
# mental_illness <- readRDS(file = here::here("models/mental_illness.rds"))
# 
# modelsummary(
#   list(
#     "Mental Illness Exist" = mental_illness
#   ),
#   statistic = "mad"
#   )
```

@@

# Additional data details {#sec-appendix-data-details}

## Variable Definitions
Variable: "AGE"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">AGE</td></tr><tr><td class="label">Label:</td><td class="value">Age</td></tr><tr><td class="label">Variable Text:</td><td class="value">AGE reports the individual's exact age, calculated from date of birth, as of the last day (12/31) of the survey year. <br>&amp;nbsp;Date
 of birth and age were asked for each reporting unit member, and then 
exact age was calculated from date of birth. Where the calculated age 
and the age provided did not match, inconsistencies were reviewed and 
resolved. When date of birth was not provided, but age was provided, the
 month and year of birth were assigned randomly from among the possible 
valid options. For any cases still not accounted, age was imputed using:<br>(1) the mean age difference between MEPS participants with certain family relationships (where available) or  <br>(2) the mean age value for MEPS participants.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">99</td></tr><tr><td class="label">End Position:</td><td class="value">101</td></tr><tr><td class="label">Width:</td><td class="value">3</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="label">Coder Instructions:</td><td class="value">CodesAGE is a 3-digit-numeric variable.<br><br>085:	Top code for 85 years or older (2001-forward)<br>090:	Top code for 90 years or older (1996-2000)<br>996:	Missing</td></tr></tbody></table>

```
Variable: "Sex"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">SEX</td></tr><tr><td class="label">Label:</td><td class="value">Sex</td></tr><tr><td class="label">Variable Text:</td><td class="value">SEX indicates whether the person was male or female.  <br><br>Collection
 of information on SEX in MEPS&amp;nbsp;Data on the sex of each 
reporting unit (RU) member was determined during the NHIS interview, and
 was then verified, and, if necessary, corrected during each MEPS 
interview.<br><br>If the respondent was a new RU member or their sex was
 not ascertained in the NHIS interview, MEPS initially used the first 
name of the respondent to assign their sex. If the first name gave no 
clear indication of sex, the reported family relationships were used to 
assign sex. If the sex of the respondent was still unclear, sex was 
randomly assigned. <br><br>The NHIS method of ascertaining the sex of 
the respondent, which primarily informs the MEPS version of SEX, is 
similar to the MEPS method. First, sex of the respondent was inferred 
from the individual's first name or family relationships. If the sex of 
the respondent was unclear, the interviewer was instructed to explicitly
 ask the person's sex. Beginning in 1998, interviewers were told to 
"enter your best guess" when the respondent either did not know or 
refused to answer the direct question that was asked about the person's 
sex.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">102</td></tr><tr><td class="label">End Position:</td><td class="value">102</td></tr><tr><td class="label">Width:</td><td class="value">1</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">1</td><td class="cat_lbl">Male</td></tr><tr><td class="cat_val">2</td><td class="cat_lbl">Female</td></tr><tr><td class="cat_val">7</td><td class="cat_lbl">Unknown-refused</td></tr><tr><td class="cat_val">8</td><td class="cat_lbl">Unknown-not ascertained</td></tr><tr><td class="cat_val">9</td><td class="cat_lbl">Unknown-don't know</td></tr></tbody></table>
```
Variable: "MARSTAT"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">MARSTAT</td></tr><tr><td class="label">Label:</td><td class="value">Legal marital status</td></tr><tr><td class="label">Variable Text:</td><td class="value">For persons age 16 and older, MARSTAT reports the person's legal marital status.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">103</td></tr><tr><td class="label">End Position:</td><td class="value">104</td></tr><tr><td class="label">Width:</td><td class="value">2</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">00</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">10</td><td class="cat_lbl">Married</td></tr><tr><td class="cat_val">20</td><td class="cat_lbl">Widowed</td></tr><tr><td class="cat_val">30</td><td class="cat_lbl">Divorced</td></tr><tr><td class="cat_val">40</td><td class="cat_lbl">Separated</td></tr><tr><td class="cat_val">50</td><td class="cat_lbl">Never married</td></tr><tr><td class="cat_val">99</td><td class="cat_lbl">Unknown marital status</td></tr></tbody></table>
```
Variable: "RACEA"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">RACEA</td></tr><tr><td class="label">Label:</td><td class="value">Main Racial Background (Pre-1997 Revised OMB Standards), self-reported or interviewer reported</td></tr><tr><td class="label">Variable Text:</td><td class="value">RACEA
 reports the race of the respondent. If not ascertained, the race and/or
 ethnicity were assigned based on relationship to other members of the 
DU using a priority ordering that gave precedence to blood relatives in 
the immediate family.</td></tr><tr><td class="label">Concept:</td><td class="value">Ethnicity/Nativity Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">105</td></tr><tr><td class="label">End Position:</td><td class="value">107</td></tr><tr><td class="label">Width:</td><td class="value">3</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">100</td><td class="cat_lbl">White</td></tr><tr><td class="cat_val">200</td><td class="cat_lbl">Black/African-American</td></tr><tr><td class="cat_val">300</td><td class="cat_lbl">Aleut, Alaskan Native, or American Indian</td></tr><tr><td class="cat_val">310</td><td class="cat_lbl">Alaskan Native or American Indian</td></tr><tr><td class="cat_val">320</td><td class="cat_lbl">Alaskan Native/Eskimo</td></tr><tr><td class="cat_val">330</td><td class="cat_lbl">Aleut</td></tr><tr><td class="cat_val">340</td><td class="cat_lbl">American Indian</td></tr><tr><td class="cat_val">350</td><td class="cat_lbl">American Indian or Alaskan Native and any other group</td></tr><tr><td class="cat_val">400</td><td class="cat_lbl">Asian or Pacific Islander</td></tr><tr><td class="cat_val">410</td><td class="cat_lbl">Asian</td></tr><tr><td class="cat_val">411</td><td class="cat_lbl">Chinese</td></tr><tr><td class="cat_val">412</td><td class="cat_lbl">Filipino</td></tr><tr><td class="cat_val">413</td><td class="cat_lbl">Korean</td></tr><tr><td class="cat_val">414</td><td class="cat_lbl">Vietnamese</td></tr><tr><td class="cat_val">415</td><td class="cat_lbl">Japanese</td></tr><tr><td class="cat_val">416</td><td class="cat_lbl">Asian Indian</td></tr><tr><td class="cat_val">420</td><td class="cat_lbl">Pacific Islander</td></tr><tr><td class="cat_val">421</td><td class="cat_lbl">Hawaiian</td></tr><tr><td class="cat_val">422</td><td class="cat_lbl">Samoan</td></tr><tr><td class="cat_val">423</td><td class="cat_lbl">Guamanian</td></tr><tr><td class="cat_val">430</td><td class="cat_lbl">Other Asian or Pacific Islander</td></tr><tr><td class="cat_val">431</td><td class="cat_lbl">Other Asian or Pacific Islander (1992-1995)</td></tr><tr><td class="cat_val">432</td><td class="cat_lbl">Other Asian or Pacific Islander (1996)</td></tr><tr><td class="cat_val">433</td><td class="cat_lbl">Other Asian or Pacific Islander (1997-1998)</td></tr><tr><td class="cat_val">434</td><td class="cat_lbl">Other Asian (1999 forward)</td></tr><tr><td class="cat_val">500</td><td class="cat_lbl">Other Race</td></tr><tr><td class="cat_val">510</td><td class="cat_lbl">Other Race (1963-1977)</td></tr><tr><td class="cat_val">520</td><td class="cat_lbl">Other Race (1978)</td></tr><tr><td class="cat_val">530</td><td class="cat_lbl">Other Race (1979-1991)</td></tr><tr><td class="cat_val">540</td><td class="cat_lbl">Other Race (1992-1995)</td></tr><tr><td class="cat_val">550</td><td class="cat_lbl">Other Race (1996)</td></tr><tr><td class="cat_val">560</td><td class="cat_lbl">Other Race (1997-1998)</td></tr><tr><td class="cat_val">570</td><td class="cat_lbl">Other Race (1999-2002)</td></tr><tr><td class="cat_val">580</td><td class="cat_lbl">Primary Race not releasable</td></tr><tr><td class="cat_val">600</td><td class="cat_lbl">Multiple Race, No Primary Race Selected</td></tr><tr><td class="cat_val">610</td><td class="cat_lbl">Multiple Race, including Asian, excluding Black and White</td></tr><tr><td class="cat_val">611</td><td class="cat_lbl">Multiple Race, including Asian and Black, excluding White</td></tr><tr><td class="cat_val">612</td><td class="cat_lbl">Multiple Race, including Asian and White, excluding Black</td></tr><tr><td class="cat_val">613</td><td class="cat_lbl">Multiple Race, including Black, excluding Asian and White</td></tr><tr><td class="cat_val">614</td><td class="cat_lbl">Multiple Race, including Black and White, excluding Asian</td></tr><tr><td class="cat_val">615</td><td class="cat_lbl">Multiple Race, including White, excluding Asian and Black</td></tr><tr><td class="cat_val">616</td><td class="cat_lbl">Multiple Race, including Asian, White, and Black</td></tr><tr><td class="cat_val">617</td><td class="cat_lbl">Multiple Race, excluding Asian, White, and Black</td></tr><tr><td class="cat_val">900</td><td class="cat_lbl">Unknown</td></tr><tr><td class="cat_val">970</td><td class="cat_lbl">Unknown-refused</td></tr><tr><td class="cat_val">980</td><td class="cat_lbl">Unknown-not ascertained</td></tr><tr><td class="cat_val">990</td><td class="cat_lbl">Unknown (1997 forward: Don't know)</td></tr></tbody></table>
```
Variable: "MARSTAT"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">MARSTAT</td></tr><tr><td class="label">Label:</td><td class="value">Legal marital status</td></tr><tr><td class="label">Variable Text:</td><td class="value">For persons age 16 and older, MARSTAT reports the person's legal marital status.</td></tr><tr><td class="label">Concept:</td><td class="value">Core Demographic Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">103</td></tr><tr><td class="label">End Position:</td><td class="value">104</td></tr><tr><td class="label">Width:</td><td class="value">2</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">00</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">10</td><td class="cat_lbl">Married</td></tr><tr><td class="cat_val">20</td><td class="cat_lbl">Widowed</td></tr><tr><td class="cat_val">30</td><td class="cat_lbl">Divorced</td></tr><tr><td class="cat_val">40</td><td class="cat_lbl">Separated</td></tr><tr><td class="cat_val">50</td><td class="cat_lbl">Never married</td></tr><tr><td class="cat_val">99</td><td class="cat_lbl">Unknown marital status</td></tr></tbody></table>
```
Variable: "EDUC"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">EDUC</td></tr><tr><td class="label">Label:</td><td class="value">Educational attainment</td></tr><tr><td class="label">Variable Text:</td><td class="value">EDUC
 reports the highest level of schooling an individual had completed, in 
terms of completed grades for persons with less than a high school 
degree, and in terms of degrees attained for high school graduates and 
those with higher education. EDUC is available for all survey 
participants age 5 and older at the time of their first MEPS interview. 
It is an IPUMS MEPS constructed variable. <br><br>See Comparability section for more information on the construction of EDUC.<br><br>Education
 Variables offered through IPUMS MEPS&amp;nbsp;In addition to EDUC, 
there are several other educational attainment variables offered through
 MEPS. Unlike EDUC, these variables are not available for every MEPS 
year.<br><br>EDUCYR: Years of education completed (available 1996-2011 and 2014 forward)<br><br>HIDEG: Highest degree completed (available 1996-2011 and 2014 forward)<br><br>EDUYRDG: Years of education and highest degree completed (available 2011-2015)<br><br>EDRECODE: Years of education, recode (available 2011-2015)<br></td></tr><tr><td class="label">Concept:</td><td class="value">Education Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">108</td></tr><tr><td class="label">End Position:</td><td class="value">110</td></tr><tr><td class="label">Width:</td><td class="value">3</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">000</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">100</td><td class="cat_lbl">Grade 12 or less, no high school diploma or equivalent</td></tr><tr><td class="cat_val">101</td><td class="cat_lbl">Grade 8 or less (no further detail)</td></tr><tr><td class="cat_val">102</td><td class="cat_lbl">Never attended/kindergarten only</td></tr><tr><td class="cat_val">103</td><td class="cat_lbl">Grades 1-11 (no further detail)</td></tr><tr><td class="cat_val">104</td><td class="cat_lbl">Grade 1</td></tr><tr><td class="cat_val">105</td><td class="cat_lbl">Grade 2</td></tr><tr><td class="cat_val">106</td><td class="cat_lbl">Grade 3</td></tr><tr><td class="cat_val">107</td><td class="cat_lbl">Grade 4</td></tr><tr><td class="cat_val">108</td><td class="cat_lbl">Grade 5</td></tr><tr><td class="cat_val">109</td><td class="cat_lbl">Grade 6</td></tr><tr><td class="cat_val">110</td><td class="cat_lbl">Grade 7</td></tr><tr><td class="cat_val">111</td><td class="cat_lbl">Grade 8</td></tr><tr><td class="cat_val">112</td><td class="cat_lbl">Grade 9-12, no diploma (no further detail)</td></tr><tr><td class="cat_val">113</td><td class="cat_lbl">Grade 9</td></tr><tr><td class="cat_val">114</td><td class="cat_lbl">Grade 10</td></tr><tr><td class="cat_val">115</td><td class="cat_lbl">Grade 11</td></tr><tr><td class="cat_val">116</td><td class="cat_lbl">12th grade, no diploma</td></tr><tr><td class="cat_val">200</td><td class="cat_lbl">High school diploma or GED</td></tr><tr><td class="cat_val">201</td><td class="cat_lbl">High school graduate</td></tr><tr><td class="cat_val">202</td><td class="cat_lbl">GED or equivalent</td></tr><tr><td class="cat_val">300</td><td class="cat_lbl">Some college, no 4yr degree</td></tr><tr><td class="cat_val">301</td><td class="cat_lbl">Some college, no degree</td></tr><tr><td class="cat_val">302</td><td class="cat_lbl">AA degree: technical/vocational/occupational</td></tr><tr><td class="cat_val">303</td><td class="cat_lbl">AA degree: academic program</td></tr><tr><td class="cat_val">400</td><td class="cat_lbl">Bachelor's degree (BA,AB,BS,BBA)</td></tr><tr><td class="cat_val">500</td><td class="cat_lbl">Master's, Professional, or Doctoral Degree</td></tr><tr><td class="cat_val">501</td><td class="cat_lbl">Master's degree (MA,MS,Med,MBA)</td></tr><tr><td class="cat_val">502</td><td class="cat_lbl">Professional (MD,DDS,DVM,JD)</td></tr><tr><td class="cat_val">503</td><td class="cat_lbl">Doctoral degree (PhD, EdD)</td></tr><tr><td class="cat_val">504</td><td class="cat_lbl">Other degree</td></tr><tr><td class="cat_val">505</td><td class="cat_lbl">Professional School or Doctoral degree, topcoded (MD, DDS, DVM, JD, PhD, EdD)</td></tr><tr><td class="cat_val">996</td><td class="cat_lbl">No degree, years of education unknown</td></tr><tr><td class="cat_val">997</td><td class="cat_lbl">Unknown--refused</td></tr><tr><td class="cat_val">998</td><td class="cat_lbl">Unknown--not ascertained</td></tr><tr><td class="cat_val">999</td><td class="cat_lbl">Unknown--don't know</td></tr></tbody></table>
```
Variable: "INCTOT"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">INCTOT</td></tr><tr><td class="label">Label:</td><td class="value">Total personal income</td></tr><tr><td class="label">Variable Text:</td><td class="value">INCTOT
 reports the sum of all person-level income for the current calendar 
year, excluding income from tax refunds and capital gains.<br><br>INCTOT
 includes annual earnings from wages, salaries, bonuses, tips, 
commissions; business and farm games and losses; unemployment and 
workers' compensation; interest and dividends; alimony, child support, 
and other private cash transfers; private pensions, IRA withdrawals, 
social security, and veterans payments; supplemental security income and
 cash welfare payments from public assistance, Temporary Assistance for 
Needy Families, and related programs; gains or losses from estates, 
trusts, partnerships, S corporations, rent, and royalties; and a small 
amount of "other" income. Person-level income excluded tax refunds and 
capital gains. <br><br>Logical editing or weighted, sequential hot-deck 
imputation was used to impute income amounts for missing values (both 
for item non-response and for person in the full-year file who were not 
in round 3). Reported income components were generally left unedited. <br><br>Related Variables<br>INCWAGE: annual wage and salary income of individuals<br><br>INCBUS: annual business income of individuals<br><br>INCUNEMP: annual unemployment compensation income of individuals<br><br>INCWKCOM: annual workers' compensation income of individuals<br><br>INCINT: annual interest income of individuals<br><br>INCDIVID: annual dividend income of individuals<br><br>INCRETIR: annual pension income of individuals<br><br>INCSS: annual social security income of individuals<br><br>INCTRST: annual trust/rent income of individuals<br><br>INCVET: annual veteran income of individuals<br><br>INCIRA: annual IRA income of individuals<br><br>INCALIM: annual alimony income of individuals<br><br>INCCHLD: annual child support income of individuals<br><br>INCCASH: annual other regular cash income income of individuals<br><br>INCSSI: annual SSI income of individuals<br><br>INCWELFR: annual public assistance income of individuals<br><br>INCOTH: annual other income of individuals<br></td></tr><tr><td class="label">Concept:</td><td class="value">Total Income and Earnings Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">111</td></tr><tr><td class="label">End Position:</td><td class="value">119</td></tr><tr><td class="label">Width:</td><td class="value">9</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">2</td></tr><tr><td class="label">Coder Instructions:</td><td class="value">CodesINCTOT
 is a 9-digit numeric variable with 2 implied decimals. That is, values 
of 012345678 should be interpreted as 123456.78. The command files 
delivered with IPUMS extracts automatically divide INCTOT by 100, so no 
further adjustment is needed.<br><br>9999999.96:	Not in Universe<br>9999999.97:	Unknown-refused<br>9999999.98:	Unknown-not ascertained<br>9999999.99:	Unknown-don't know</td></tr></tbody></table>
```
Variable: "K6SUM"
```{=html}
<table class="data"><tbody><tr><td class="label">Name:</td><td class="value">K6SUM</td></tr><tr><td class="label">Label:</td><td class="value">K6 score for nonspecific psychological distress: last 30 days</td></tr><tr><td class="label">Variable Text:</td><td class="value">For
 persons eligible for the self-administered questionnaire (SAQELIG), 
K6SUM is the summed scale value measuring nonspecific psychological 
distress over a 30-day recall period. This scale, developed by Ronald C.
 Kessler and known as the Kessler 6 Scale (K6), asks about six 
manifestations of nonspecific psychological distress.<br><br>Kessler 
recommends scoring the scale by assigning 0 to 4 points for each of the 
six questions, based on the reported frequency of the feelings (i.e., 0 
for "none of the time"; 1 for "a little of the time"; 2 for "some of the
 time"; 3 for "most of the time"; and 4 for "all of the time"). The 
range for summed responses on the K6 Scale is thus 0 to 24, with 0 
suggesting the lowest level of nonspecific psychological distress, and 
24 suggesting the highest level of nonspecific psychological distress. 
According to the scoring criteria proposed by Kessler, persons with a 
score of 13 or greater are likely to be experiencing severe mental 
illness.<br><br>Kessler's instrument asks how often, during the past 30 days, the respondent felt:<br><br>So sad that nothing could cheer you up? (ASAD)<br><br>Nervous? (ANERVOUS)<br><br>Restless or fidgety? (ARESTLESS)<br><br>Hopeless? (AHOPELESS)<br><br>That everything was an effort? (AEFFORT)<br><br>Worthless? (AWORTHLESS)<br>As noted above, acceptable responses fell into five categories, ranging from "none of the time" to "all of the time."</td></tr><tr><td class="label">Concept:</td><td class="value">Adult Mental Health Variables -- PERSON</td></tr><tr><td class="label">Start Position:</td><td class="value">120</td></tr><tr><td class="label">End Position:</td><td class="value">121</td></tr><tr><td class="label">Width:</td><td class="value">2</td></tr><tr><td class="label">Variable Format:</td><td class="value">numeric</td></tr><tr><td class="label">Implied Decimal Places:</td><td class="value">0</td></tr><tr><td class="table_heading" colspan="2">Categories</td></tr><tr><td colspan="2" class="value"></td></tr></tbody></table>
<table class="cats"><tbody><tr><th class="cat_val">Value</th><th class="cat_lbl">Label</th></tr><tr><td class="cat_val">96</td><td class="cat_lbl">NIU</td></tr><tr><td class="cat_val">98</td><td class="cat_lbl">Unknown-not ascertained</td></tr></tbody></table>
```
